---
title: "Lab Week 5, Time Series"
author: "Melissa Gordon Wolf"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning=FALSE,
                      message=FALSE)
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble) #time series tibbles
library(fable)#time series forecasting
library(feasts)#time series exploration
library(fabletools)
library(forecast)
library(sf)
library(tmap)
library(mapview)
library(sjmisc)
```

## Monthly US energy consumption (renewables)

```{r}
us_renew <- read_csv(here("data","renewables_cons_prod.csv")) %>% 
  clean_names()  #I think this just makes it lowercase
```

```{r results='hide'}
unique(us_renew$description)  #gives you number of unique values
```

- Make the descriptoin all lowercase
- Only keep observations where the "description" variable contains the string pattern "consumption" (to only keep the consumption data)
  - Remove any observations where "description" variable contains "total"
  
```{r}
renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% #overwrite the existing column so that it only contains lowercase letters
  filter(str_detect(description,pattern="consumption")) %>% #looks for partial matches that contains the string lowercase consumption and keep it, remove all others
  filter(!str_detect(description, pattern="total")) #remove everything with total

frq(renew_clean$description)  #check
```

### Convert `yyyymm` column to a date

```{r}
renew_date <- renew_clean %>% 
  mutate(yr_mo_day=lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep=tsibble::yearmonth(yr_mo_day)) %>% #only pull the year and month
  mutate(value=as.numeric(value)) %>% #just change to numeric
  drop_na(month_sep,value)
  
class(renew_date$month_sep)

#parse data using lubridate::month
renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day,label=TRUE)) %>% 
  mutate(year=year(yr_mo_day))

#490 failed to parse means it doesn't like the 490 NA's but it's still okay and worked
```

### Look at it:

```{r}
rewnew_gg <- ggplot(data=renew_date, aes(x=month_sep,
                                         y=value,
                                         group=description))+ #don't need this here because we have it below, but adding it anyway for pedagogy
  geom_line(aes(color=description))

rewnew_gg

View(palettes_d_names) #we need 7 colors, so we need a palette with at least 7 colors
```

Update colors with paletteer palettes:

```{r}
rewnew_gg+
  scale_color_paletteer_d("miscpalettes::sunset")
```

Convert an entire dataframe to a tsibble so that you can use feast and fable on it (for modeling and forecasting)

```{r}
renew_ts <- as_tsibble(renew_parsed,key=description, index=month_sep)

#index is tsibble compatible time variable that we created, which is month_sep
```

Let's look at our time series data in a couple of different ways:
```{r}
renew_ts %>% autoplot(value)

#autoplot does a pretty good job of creating a ggplot

renew_ts %>% gg_subseries(value)

#what each value looks like across months

renew_ts %>% gg_season(value)

#to reproduce gg_season using ggplot:
#ggplot(data=renew_parsed, aes(x=month,y=value,group=year))+
#  geom_line(aes(color=year))+
#  facet_wrap(~description,
#             ncol=1,
#             scales="free",
#             strip.position = "right")
```


### just look at the hydroelectric energy consumption
```{r}
hydro_ts <- renew_ts %>% 
  filter(description=="hydroelectric power consumption")

hydro_ts %>% autoplot(value)

hydro_ts %>% gg_season(value)
```


### what if I want the quarterly average consumption for hydro?

index by is specifically for tsibble data (instead of group_by)

```{r}
hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu=~(yearquarter(.))) %>%  # . = based on existing
  summarize(avg_consumption=mean(value))
head(hydro_quarterly)
```

### Decompose that hydro_ts

```{r}
dcmp <- hydro_ts %>% 
  model(feasts::STL(value~season(window=5)))

components(dcmp) %>% autoplot()

#hist(components(dcmp$remainder))

#1. actual value of obse
#2. moving average trend
#3. season component
#4. residual
```

### Now, look at the ACF:

This means we have some repetition in how we consume hydroelectric pattern

```{r}
hydro_ts %>% 
  ACF(value) %>% 
  autoplot()
```

### Now, model (forecasting using fable)

Be careful with this

```{r}
hydro_model <- hydro_ts %>%
  model(
    ARIMA(value),
    ETS(value)
  ) %>% 
  fabletools::forecast(h="4 years")

#don't just run the model to fit the data, but ask fable to predict

#ETS gives you another type of model and then you can compare their forecasting

#run a bunch of conservative models and see if they're all lining up

hydro_model %>% autoplot() #only giving us the forecasted values

hydro_model %>% autoplot(filter(hydro_ts, year(month_sep)>2010))

#built into arima forecasting: more uncertainty as we move towards the future
#think really hard about how valid this might be and if the model you're using is really representing your data

```

### Now, let's make a world map!

```{r}
world <- read_sf(dsn=here("data","TM_WORLD_BORDERS_SIMPL-0.3-1"), #tell it what directory to find it in
                 layer="TM_WORLD_BORDERS_SIMPL-0.3") 
mapview(world)
```

